Quantifying maternal reproductive output of chondrichthyan fishes

### 2024-09-29 README.qmd
readme

Harry AV, Baremore IE & Piercy AN (2024) Quantifying maternal reproductive output of chondrichthyan fishes. *Canadian Journal of Fisheries and Aquatic Sciences*

@@

Repository structure

fs::dir_tree()

@@

```{r}
data %>% 
  ggplot() + 
  geom_histogram(aes(x = x, fill = as.factor(z)), binwidth = 5, col = "black") +
  labs(x = "Fork length (cm)", y = "Proportion in maternal condition") + 
  guides(fill = guide_legend(title = "Maternal stage")) + 
  theme(legend.position = "top")

@@

@@

ggplot()  + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  ylim(0, 1) +
  labs(x = "Fork length (cm)", y = "Proportion in maternal condition")
```

@@

```{r message=FALSE, warning=FALSE}
library(TMB)
compile("code/logistic3.cpp")
dyn.load(dynlib("code/logistic3"))
```

Initialize a list of parameters for the model with their starting values, then construct the objective function to be minimised using R.

```{r}
# Parameter list
pars <- list(m50 = 160, m95 = 170, c = 0.5)

@@

>>>>

Harry AV, Baremore IE & Piercy AN (2024) Quantifying maternal reproductive output of chondrichthyan fishes. *Canadian Journal of Fisheries and Aquatic Sciences.* <https://doi.org/10.1139/cjfas-2024-0031>

@@

Repository structure:

```{r, include = TRUE, eval=TRUE, echo=FALSE}
fs::dir_tree()
```

@@

```{r, include = TRUE, eval=TRUE, echo=FALSE}
data %>% 
  filter(!is.na(y)) %>% 
  ggplot() + 
  geom_histogram(aes(x = x, fill = as.factor(y)), binwidth = 5, col = "black") +
  labs(x = "Fork length (cm)", y = "Proportion in mature condition", title = "Length distribution showing female maturity stage") + 
  guides(fill = guide_legend(title = "Maturity stage")) + 
  theme(legend.position = c(0.2, 0.75))
```

```{r, include = TRUE, eval=TRUE, echo=FALSE}
data %>% 
  ggplot() + 
  geom_histogram(aes(x = x, fill = as.factor(z)), binwidth = 5, col = "black") +
  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length distribution showing female maternity stage") + 
  guides(fill = guide_legend(title = "Maternal stage")) + 
  theme(legend.position = c(0.2, 0.75))

@@

```{r message=FALSE, warning=FALSE, include=TRUE}
brks = seq(40, 220, 10)
data_binned <- data %>%
  filter(!is.na(y)) %>% 
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(y)/ n())

ggplot()  + 
  geom_rug(data = filter(data, y == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, y == 0), aes(x = x), sides = "b") +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  ylim(0, 1) +
  labs(x = "Fork length (cm)", y = "Proportion in mature condition", title = "Length at maturity data for female sandbar sharks") 
```

@@

ggplot()  + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  ylim(0, 1) +
  labs(x = "Fork length (cm)", y = "Proportion in maternal condition", title = "Length at maternity data for female sandbar sharks") 
```

@@

```{r message=FALSE, warning=FALSE, results="hide"}
library(TMB)
compile("code/logistic3.cpp")
dyn.load(dynlib("code/logistic3"))
```

Initialize a list of parameters for the model with their starting values, then construct the objective function to be minimised using R. Next use `nlminb` to optimize the model. `nlminb` takes as its arguments a vector of the parameters to be estimated, the function to be optimized, an analytical gradient vector that aids with optimization, and vectors of maximum and minimum allowable values for the model parameters.

```{r message=FALSE, warning=FALSE, results = "hide"}
# Parameter list
pars <- list(m50 = 160, m95 = 170, c = 0.5)

@@

To check whether the model has successfully converged, run the following:

```{r}
opt$convergence
```

The maximum likelihood values of parameters estimated by `TMB` can be obtained as follows:

```{r }
opt$par
```

The estimated length at 50% maternity ($L_{50}$) and length at 95% maternity ($L_{95}$) are 160 and 174cm, respectively. $L_{50}$ is estimated length at which half of the females in the population were in maternal condition, meaning they would have given birth and contributed to recruitment within that year. The final parameter is the asymptotic of the logistic function, $P_{Max}$, which is the maximum proportion of females in a given year in maternal condition, in this case 0.48 (or 48% of the population).
